---
title: "Week 2 Homework airquality"
author: "Rachel Saidi"
date: "July 31, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load in the Dataset. 

Because airquality is a pre-built dataset, we can write it to our data directory to store it for later use.

```{r include = FALSE }
library(tidyverse)
datasets::airquality
setwd("C:/Users/rsaidi/Dropbox/Rachel/MontColl/Datasets/Datasets")
write_csv(airquality, "airquality.csv")
```

# Look at the structure of the data

```{r}
str(airquality)
```

# Change the Months from 5 - 9 to May through September

```{r}
airquality$Month[airquality$Month == 5]<- "May"
airquality$Month[airquality$Month == 6]<- "June"
airquality$Month[airquality$Month == 7]<- "July"
airquality$Month[airquality$Month == 8]<- "August"
airquality$Month[airquality$Month == 9]<- "September"
```


# Look at the summary statistics of the dataset

```{r}
str(airquality)
summary(airquality)
```

If you want to look at specific statistics, here are some variations on coding

```{r}
mean(airquality$Temp)
mean(airquality[,4])
median(airquality$Temp)
var(airquality$Wind)
```

# Reorder the Months so they do not default to alphabetical

```{r}
airquality$Month<-factor(airquality$Month, levels=c("May", "June","July", "August", "September"))
```



#   Plot 1:  Create a histogram categorized by Month with qplot

```{r}
p1 <- qplot(data = airquality,Temp,fill = Month,geom = "histogram", bins = 20)
p1
```

# Plot 2: Make a histogram using ggplot (instead of qplot)

Reorder the legend so that it is not the default (alphabetical), but rather in order that months come

```{r}
p2 <- airquality %>%
  ggplot(aes(x=Temp, fill=Month)) +
  geom_histogram(position="identity", alpha=0.5, binwidth = 5, color = "white")+
  scale_fill_discrete(name = "Month", labels = c("May", "June","July", "August", "September"))
p2
```


# Plot 3: Create side-by-side boxplots categorized by Month

```{r}
p3 <- airquality %>%
  ggplot(aes(Month, Temp, fill = Month)) + 
  ggtitle("Temperatures") +
  xlab("Months") +
  ylab("Frequency") +
  geom_boxplot() +
  scale_fill_discrete(name = "Month", labels = c("May", "June","July", "August", "September"))
p3 
```


# Plot 4: Make the same side-by-side boxplots, but in grey-scale

```{r}
p4 <- airquality %>%
  ggplot(aes(Month, Temp, fill = Month)) + 
  ggtitle("Temperatures") +
  xlab("Temperatures") +
  ylab("Frequency") +
  geom_boxplot()+
  scale_fill_grey(name = "Month", labels = c("May", "June","July", "August", "September"))
p4
```



