---
title: "DATA 110 Week 2 Homework airquality plots"
author: "Steve Dutky"
date: "September 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Airquality dataset included with tidyverse
```{r}
library(tidyverse)
airquality<- datasets::airquality
```
## see what the included airquality dataset looks like:

```{r}
# structure
str(airquality)
```

```{r}
#
# variety
summary(airquality)
```

## remap months from integers to names
```{r}
airquality$Month<-as.factor(airquality$Month)
levels(airquality$Month)=c("May","Jun","Jul","Aug","Sep")
```

# Plot 1: Create a histogram categorized by Month with qplot
```{r}
p1 <- qplot(data = airquality,Temp,fill = Month,geom = "histogram", bins = 20)
p1
```

# Plot 2: Make a histogram using ggplot (instead of qplot)

```{r}
p2 <- airquality %>%
  ggplot(aes(x=Temp, fill=Month)) +
  geom_histogram(position="identity", alpha=1, binwidth = 5, color = "white")+
  scale_fill_discrete(name = "Month",  labels = levels(airquality$Month))
p2
```



# Plot 3: Create side-by-side boxplots categorized by Month
## y-axis appears to be temperatures, not frequencies

```{r}
p3 <- airquality %>%
  ggplot(aes(Month, Temp, fill = Month)) + 
  ggtitle("Temperatures") +
  xlab("Months") +
  ylab("Frequency") +
  geom_boxplot() +
  scale_fill_discrete(name = "Month", labels = levels(airquality$Month))
p3 
```


# Plot 4: Make the same side-by-side boxplots, but in grey-scale
```{r}
p4 <- airquality %>%
  ggplot(aes(Month, Temp, fill = Month)) + 
  ggtitle("Temperatures") +
  xlab("Temperatures") +
  ylab("Frequency") +
  geom_boxplot()+
  scale_fill_grey(name = "Month", labels = levels(airquality$Month))
p4
```


# **My turn:**
##
# **Plot 2a: Make a barplot using ggplot using fill color for Temperature**
## How to do this with continuous Temp instead of factor?
## Had to use negative Temp to get temperatures to rise from bottom

```{r}
airq<-airquality
airq$Temp<-as.factor(-airq$Temp)

p2a <- airq %>%
  ggplot(aes(x=Month, fill=(Temp))) +
  geom_bar(stat = "count")+
  scale_fill_discrete(name = "Temp",  labels =levels(airq$Temp))
p2a
```

# **Plot 3a: Create temperature heat map categorized by Month**

```{r}
p3a <- airquality %>%
  ggplot(aes(Month,Temp)) + 
  ggtitle("Temperatures") +
  xlab("Months") +
  ylab("Temperatures") +
  geom_tile(aes(fill=Temp)) + 
  scale_fill_gradient2(low="blue",mid="purple",high = "red",midpoint = median(airquality$Temp),limits=range(airquality$Temp))
p3a 
```

# **Plot 3b: Another try from Stackoverflow:**
## works better than the other tries
```{r}
p3b<-airquality %>%
  ggplot(aes(Month,group=-Temp,fill=Temp))+
  geom_bar()+
  scale_fill_continuous(low="blue",high="red")
p3b  
```

# **[My GitHub account](https://github.com/sdutky/mcData110)**
## https://github.com/sdutky/mcData110